---
title: "Dial in the model"
author: "Michael Koontz"
date: "4/24/2018"
output: html_document
---

```{r setup, message = FALSE}
library(lubridate)
library(sf)
library(lme4)
library(dplyr)
library(ggplot2)
library(here)
library(lubridate)
library(multcomp)
```

```{r read_files, message = FALSE}
fires <- st_read(dsn = here::here("/data/features/fire_perim/fire_perim_sn_16_1_shp/"),
                 stringsAsFactors = FALSE) %>% 
  st_transform(4326)
sn <- st_read(here::here("/data/features/SierraEcoregion_Jepson/SierraEcoregion_Jepson.shp")) %>% 
  st_transform(4326)

if (!file.exists(here::here("data/data_output/all-fire-samples_texture.rds"))) {
  source(here::here("data/data_carpentry/merge_fire_samples.R"))
}

load(here::here("data/data_output/all-fire-samples_texture.rds"))

samps <- texture_samps
```

```{r look_at_samples, cache = FALSE}
# Extract unique Fire ID from the sample point IDs
samps$fire_id <- substr(as.character(samps$id), start = 1, stop = 20)
```

## Topography variables

Adjust aspect to reflect "heat load index" [@McCune2002] with southwest having a value of 1, and northeast having a value of -1

```{r adjust_aspect}
circular_aspect <- function(aspect) {
  
  new_aspect <- ((-1 * (aspect)) * pi / 180) - (3/4 * pi)
  return(cos(new_aspect))
  
}

samps$c_aspect <- circular_aspect(samps$aspect)
```


Calculate predicted potential annual direct incident radiation (also from @McCune2002) which combines slope, aspect, and latitude

```{r calc_heat_load}
fold_aspect <- function(asp) {
  
  rad <- asp * pi / 180
  folded_aspect <- pi - abs(rad - pi)
  return(folded_aspect)
  
}

samps$folded_aspect <- fold_aspect(samps$aspect)

# G column: latitude
# H column: slope
# I column: folded aspect

# Potential 
pdir <- function(lat, slope, folded_asp) {
  lat <- lat * pi / 180
  slope <- slope * pi / 180
  exp(-1.467 + 1.582 * cos(lat) * cos(slope) - 1.5 * cos(folded_asp) * sin(slope) * sin(lat) - 0.262 * sin(lat) * sin(slope) + 0.607 * sin(folded_asp) * sin(slope))
}

samps$pdir <- pdir(samps$lat, samps$slope, samps$folded_aspect)
```

Consolidate the names of the topography variables
```{r consolidate_topo_vars}
topo_vars <- c("c_aspect", "pdir", "elev", "lat", "lon", "slope", "topo_roughness_1", "topo_roughness_2", "topo_roughness_3", "topo_roughness_4")
```

## Timing variables

Adjust ordinal day to reflect "middle of summer-ness" with August 6th (the 218th day of the year) getting a 1 and February 4th (the 35th day of the year) getting a value of -1

```{r adjust_ordinal_day}
circular_doy <- function(doy) {
  
  new_doy <- ((doy - 218) / 365) * 2 * pi
  return(cos(new_doy))
  
}

samps$c_doy <- circular_doy(samps$ordinal_day)
```

Make year an integer

```{r year_to_integer}
samps$year_ <- as.numeric(as.character(samps$year_))
```

Consolidate timing variable names.

```{r consolidate_timing_vars}
timing_vars <- c("year_", "alarm_date", "cont_date", "c_doy")
```

## Vegetation variables

Consolidate vegetation variable names.

```{r consolidate_veg_var_names}
veg_vars <-
  sapply(X = 1:4, 
         FUN = function(i) paste(
           c(
             "focal_mean_ndvi",
             "focal_mean_ndwi",
             "het_ndvi",
             "het_ndwi",
             "ndvi_asm",
             "ndvi_contrast",
             "ndvi_corr",
             "ndvi_dent",
             "ndvi_diss",
             "ndvi_dvar",
             "ndvi_ent",
             "ndvi_idm",
             "ndvi_imcorr1",
             "ndvi_imcorr2",
             "ndvi_inertia",
             "ndvi_prom",
             "ndvi_savg",
             "ndvi_sent",
             "ndvi_shade",
             "ndvi_svar",
             "ndvi_var"
           ), i, sep = "_")) %>% 
  as.vector() %>% 
  c("preFire_ndvi", "preFire_ndwi")
```

## Fire weather variables

Consolidate fire weather variable names.

```{r consolidate_fire_weather_var_names}
fireWeather_vars <- c("erc", "fm100", "tmmx")
```

# Scale all predictor variables

First consolidate all variable names.

```{r consolidate_var_names}
all_vars <- c(topo_vars, timing_vars, veg_vars, fireWeather_vars)
```

Just focus on the points from the mixed conifer/yellow pine forest

```{r conifer_subset}
mixed_con <- 
  samps %>%
  filter(conifer_forest == 1)
```

Scale the predictors (ss = scaled samples)

```{r scale_samples}
ss <-
  mixed_con %>%
  mutate_at(.vars = all_vars, .funs = funs(s = as.numeric(scale(.))))

glimpse(ss)  
```

## High severity/non-high severity

```{r get_cbi_calibration_model_summary}
model_summary <- read.csv(here::here("data/data_output/cbi_calibration_model_comparison.csv"), stringsAsFactors = FALSE)
target_model <- model_summary[model_summary$response == "RBR" &
                                model_summary$time_window == 48 &
                                model_summary$interpolation == "bicubic", ]
```

Here are the other thresholds for the model using bicubic interpolation and a 48 day window:

- `r target_model$unchanged` equates to a CBI of 0
- `r target_model$low_sev` equates to a CBI of 0.1 -- threshold between "unchanged" and "low"
- `r target_model$mod_sev` equates to a CBI of 1.25 -- threshold between "low" and "medium"
- `r target_model$hi_sev` equates to a CBI of 2.25 -- threshold between "medium" and "high"

```{r high_v_not-high}
ss$stand_replacing <- ifelse(ss$RBR > target_model$hi_sev, yes = 1, no = 0)
```

Transform RBR to linear scale CBI

For 100-hour fuel moisture, we refer to @Stephens2012 who find 7.7% to be the 80th
percentile condition, 6.6% to be the 90th percentile condition, and 4.2% to be 
the 97.5th percentile condition. See also the note in @Stephens2013a about the significance
of 80th percentile conditions.

Create a variable for "extreme conditions" versus "non-extreme" conditions (with
respect to fuel moisture) and interact it with the heterogeneity variable.

```{r extreme_fm100}
# Extreme percentiles correspeond to 80th, 90th, and 97.5th percentiles of 100 hour
# fuel moisture
extreme_fm100_percentiles <- c(7.7, 6.6, 4.2)
ss$extreme80_fm100 <- ifelse(ss$fm100 < 7.7, yes = 1, no = 0)
ss$extreme90_fm100 <- ifelse(ss$fm100 < 6.6, yes = 1, no = 0)
ss$extreme97.5_fm100 <- ifelse(ss$fm100 < 4.2, yes = 1, no = 0)
```

Consider also subsetting to only places that burned to avoid any impact of unburned area that happens to be within the fire perimeter.

```{r conditional_on_burning}
ss_burned <- ss %>% filter(ss$RBR > target_model$low_sev)
```

```{r}
remoteSev_to_cbi <- function(data, response, a, b, c) {
  log((data[, response, drop = TRUE] - a) / b) / c
}

ss_burned$cbi <- remoteSev_to_cbi(data = ss_burned, response = "RBR", a = target_model$a, b = target_model$b, c = target_model$c)
```


## Modeling

```{r continuous_response}
m1a <- lm(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s, data = ss)
m2a <- lm(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s, data = ss)
m3a <- lm(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s, data = ss)
m4a <- lm(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s, data = ss)

summary(m1a)
summary(m2a)
summary(m3a)
summary(m4a)

m5a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m6a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m7a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m8a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m5a)
summary(m6a)
summary(m7a)
summary(m8a)

m9a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m10a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m11a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m12a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m9a)
summary(m10a)
summary(m11a)
summary(m12a)

m13a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m14a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m15a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m16a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m13a)
summary(m14a)
summary(m15a)
summary(m16a)

m17a <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m18a <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m19a <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m20a <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m17a)
summary(m18a)
summary(m19a)
summary(m20a)

AIC(m17a, m18a, m19a, m20a)

AIC(m1a, m2a, m3a, m4a, m5a, m6a, m7a, m8a, m9a, m10a, m11a, m12a, m13a, m14a, m15a, m16a, m17a, m18a, m19a, m20a) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)
```
```{r}
m1b <- lm(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s, data = ss_burned)
m2b <- lm(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s, data = ss_burned)
m3b <- lm(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s, data = ss_burned)
m4b <- lm(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s, data = ss_burned)

summary(m1a)
summary(m2a)
summary(m3a)
summary(m4a)

m5b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m6b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m7b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m8b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m5b)
summary(m6b)
summary(m7b)
summary(m8b)

m9b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m10b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m11b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m12b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m9b)
summary(m10b)
summary(m11b)
summary(m12b)

m13b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m14b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m15b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m16b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m13b)
summary(m14b)
summary(m15b)
summary(m16b)

m17b <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m18b <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m19b <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m20b <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m17b)
summary(m18b)
summary(m19b)
summary(m20b)

AIC(m17b, m18b, m19b, m20b)


AIC(m1b, m2b, m3b, m4b, m5b, m6b, m7b, m8b, m9b, m10b, m11b, m12b, m13b, m14b, m15b, m16b) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)
```


```{r binomial_response}

# Severe or not as bernoulli response, fm100 covariate, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_rawFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, fm100 covariate, only samples that were burned
m_sevOrNot_rawFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

###

# Severe or not as bernoulli response, extreme or not dummy covariate, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_extremeFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, only samples that were burned
m_sevOrNot_extremeFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_extremeFm100AndRawFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned
m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

AIC(m_sevOrNot_rawFm100_1_ss, m_sevOrNot_rawFm100_2_ss, m_sevOrNot_rawFm100_3_ss, m_sevOrNot_rawFm100_4_ss,
    m_sevOrNot_extremeFm100_1_ss, m_sevOrNot_extremeFm100_2_ss, m_sevOrNot_extremeFm100_3_ss, m_sevOrNot_extremeFm100_4_ss,
    m_sevOrNot_extremeFm100AndRawFm100_1_ss, m_sevOrNot_extremeFm100AndRawFm100_2_ss, m_sevOrNot_extremeFm100AndRawFm100_3_ss, m_sevOrNot_extremeFm100AndRawFm100_4_ss)  %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)

AIC(m_sevOrNot_rawFm100_1_ssBurned, m_sevOrNot_rawFm100_2_ssBurned, m_sevOrNot_rawFm100_3_ssBurned, m_sevOrNot_rawFm100_4_ssBurned,
    m_sevOrNot_extremeFm100_1_ssBurned, m_sevOrNot_extremeFm100_2_ssBurned, m_sevOrNot_extremeFm100_3_ssBurned, m_sevOrNot_extremeFm100_4_ssBurned,
    m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)


# Best binomial response model; all samps
summary(m_sevOrNot_extremeFm100AndRawFm100_2_ss)

# Best binomial response model; burned samps
summary(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)


m_nonextreme <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = subset(ss_burned, subset = ss_burned$extreme80_fm100 == 0), control = glmerControl(optimizer = "bobyqa"))
m_extreme <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = subset(ss_burned, subset = ss_burned$extreme80_fm100 == 1), control = glmerControl(optimizer = "bobyqa"))

summary(m_nonextreme)
summary(m_extreme)

```

```{r rbr_response}

# RBR response, fm100 covariate, all samples from within fire perimeter (some may not have burned)
m_rbr_rawFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, fm100 covariate, only samples that were burned
m_rbr_rawFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

###

# RBR response, extreme or not dummy covariate, all samples from within fire perimeter (some may not have burned)
m_rbr_extremeFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, extreme or not dummy covariate, only samples that were burned
m_rbr_extremeFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

# RBR response, extreme or not dummy covariate, interaction with raw fm100, all samples from within fire perimeter (some may not have burned)
m_rbr_extremeFm100AndRawFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned
m_rbr_extremeFm100AndRawFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

AIC(m_rbr_rawFm100_1_ss, m_rbr_rawFm100_2_ss, m_rbr_rawFm100_3_ss, m_rbr_rawFm100_4_ss,
    m_rbr_extremeFm100_1_ss, m_rbr_extremeFm100_2_ss, m_rbr_extremeFm100_3_ss, m_rbr_extremeFm100_4_ss,
    m_rbr_extremeFm100AndRawFm100_1_ss, m_rbr_extremeFm100AndRawFm100_2_ss, m_rbr_extremeFm100AndRawFm100_3_ss, m_rbr_extremeFm100AndRawFm100_4_ss)  %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)

AIC(m_rbr_rawFm100_1_ssBurned, m_rbr_rawFm100_2_ssBurned, m_rbr_rawFm100_3_ssBurned, m_rbr_rawFm100_4_ssBurned,
    m_rbr_extremeFm100_1_ssBurned, m_rbr_extremeFm100_2_ssBurned, m_rbr_extremeFm100_3_ssBurned, m_rbr_extremeFm100_4_ssBurned,
    m_rbr_extremeFm100AndRawFm100_1_ssBurned, m_rbr_extremeFm100AndRawFm100_2_ssBurned, m_rbr_extremeFm100AndRawFm100_3_ssBurned, m_rbr_extremeFm100AndRawFm100_4_ssBurned) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)


# All samps best model
summary(m_rbr_rawFm100_4_ss)

# Only burned best model
summary(m_rbr_extremeFm100_4_ssBurned)

# Model to match best model from stand replacing binomial response
summary(m_rbr_extremeFm100AndRawFm100_1_ss)
summary(m_rbr_extremeFm100AndRawFm100_2_ss)
summary(m_rbr_extremeFm100AndRawFm100_3_ss)
summary(m_rbr_extremeFm100AndRawFm100_4_ss)

```

```{r}
# Calculating the differences between B11, B12, and B13:
H0s <- rbind(het_normal = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0),
             het_xtreme = c(0, 1, 1, 0, 0, 0, 0, 1, 0, 0, 0),
             norm_xtreme = c(0, 0, -1, 0, 0, 0, 0, -1, 0, 0, 0))

# Binomial response, all samps
tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_1_ss, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ss)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_2_ss, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ss)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_3_ss, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ss)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_4_ss, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ss)

# Binomial response, burned samps
tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned)

tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, linfct = H0s)
summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)

# RBR response, all samps
tH0s_m_rbr_extremeFm100AndRawFm100_1_ss <- glht(m_rbr_extremeFm100AndRawFm100_1_ss, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_1_ss)

tH0s_m_rbr_extremeFm100AndRawFm100_2_ss <- glht(m_rbr_extremeFm100AndRawFm100_2_ss, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_2_ss)

tH0s_m_rbr_extremeFm100AndRawFm100_3_ss <- glht(m_rbr_extremeFm100AndRawFm100_3_ss, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_3_ss)

tH0s_m_rbr_extremeFm100AndRawFm100_4_ss <- glht(m_rbr_extremeFm100AndRawFm100_4_ss, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_4_ss)

# RBR response, burned samps
tH0s_m_rbr_extremeFm100AndRawFm100_1_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_1_ssBurned, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_1_ssBurned)

tH0s_m_rbr_extremeFm100AndRawFm100_2_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_2_ssBurned, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_2_ssBurned)

tH0s_m_rbr_extremeFm100AndRawFm100_3_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_3_ssBurned, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_3_ssBurned)

tH0s_m_rbr_extremeFm100AndRawFm100_4_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_4_ssBurned, linfct = H0s)
summary(tH0s_m_rbr_extremeFm100AndRawFm100_4_ssBurned)

```

```{r fit_with_brms}
library(brms)
# Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned

m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned)

m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned)

m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned)

m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned)

m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm)
# m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm)
# m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm)
# m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm)
```

Set up the proper contrasts for each model
```{r contrasts_setup}
# Contrasts
# Fix these so that fm100_s isn't 0, but rather the mean for each group (greater than 0 for non-extreme, less than 0 for extreme)

# What is the mean fm100_s for non-extreme and extreme conditions?
fm100_s_means <- 
  ss_burned %>% 
  as.data.frame() %>% 
  group_by(extreme80_fm100) %>% 
  summarize(mean_fm100 = mean(fm100_s))

het_normal <- c(0, 1, 0, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 0), mean_fm100)), 0, 0, 0, 0, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 0), mean_fm100)), 0, 0)

het_xtreme <- c(0, 1, 1, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100))
, 0, 0, 0, 1, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100))
, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100))
, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100))
)
```

Look at smallest neighborhood effect of heterogeneity:

```{r one_pixel_neighbornood}
m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm)
m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_samps[, 1:11]

het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas, 1, function(x) x %*% het_normal)
het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned)(0)

hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, breaks = 100)
ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned)(0)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned - het_xtreme_samps, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned - het_xtreme_samps)(0)
```

```{r two_pixel_neighbornood}
m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm)
m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_samps[, 1:11]

het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas, 1, function(x) x %*% het_normal)
het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned)(0)

hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, breaks = 100)
ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned)(0)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned - het_xtreme_samps, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned - het_xtreme_samps)(0)
```

```{r three_pixel_neighbornood}
m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm)
m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_samps[, 1:11]

het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas, 1, function(x) x %*% het_normal)
het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned)(0)

hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, breaks = 100)
ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned)(0)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned - het_xtreme_samps, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned - het_xtreme_samps)(0)
```

```{r four_pixel_neighbornood}
m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm)
m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_samps[, 1:11]

het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas, 1, function(x) x %*% het_normal)
het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)(0)

hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, breaks = 100)
ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)(0)

hist(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned - het_xtreme_samps, breaks = 100)
ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned - het_xtreme_samps)(0)
```


