---
title: "Dial in the model"
author: "Michael Koontz"
date: "4/24/2018"
output: html_document
---

```{r setup, message = FALSE}
library(lubridate)
library(sf)
library(lme4)
library(dplyr)
library(ggplot2)
library(here)
library(lubridate)
library(multcomp)
library(effects)
```

```{r read_files, message = FALSE}
fires <- st_read(dsn = here::here("/data/features/fire_perim/fire_perim_sn_16_1_shp/"),
                 stringsAsFactors = FALSE) %>% 
  st_transform(4326)
sn <- st_read(here::here("/data/features/SierraEcoregion_Jepson/SierraEcoregion_Jepson.shp")) %>% 
  st_transform(4326)

if (!file.exists(here::here("data/data_output/all-fire-samples_texture.rds"))) {
  source(here::here("data/data_carpentry/merge_fire_samples.R"))
}

load(here::here("data/data_output/all-fire-samples_texture.rds"))

samps <- texture_samps
```

```{r look_at_samples, cache = FALSE}
# Extract unique Fire ID from the sample point IDs
samps$fire_id <- substr(as.character(samps$id), start = 1, stop = 20)
```

## Topography variables

Adjust aspect to reflect "heat load index" [@McCune2002] with southwest having a value of 1, and northeast having a value of -1

```{r adjust_aspect}
circular_aspect <- function(aspect) {
  
  new_aspect <- ((-1 * (aspect)) * pi / 180) - (3/4 * pi)
  return(cos(new_aspect))
  
}

samps$c_aspect <- circular_aspect(samps$aspect)
```


Calculate predicted potential annual direct incident radiation (also from @McCune2002) which combines slope, aspect, and latitude

```{r calc_heat_load}
fold_aspect <- function(asp) {
  
  rad <- asp * pi / 180
  folded_aspect <- pi - abs(rad - pi)
  return(folded_aspect)
  
}

samps$folded_aspect <- fold_aspect(samps$aspect)

# G column: latitude
# H column: slope
# I column: folded aspect

# Potential 
pdir <- function(lat, slope, folded_asp) {
  lat <- lat * pi / 180
  slope <- slope * pi / 180
  exp(-1.467 + 1.582 * cos(lat) * cos(slope) - 1.5 * cos(folded_asp) * sin(slope) * sin(lat) - 0.262 * sin(lat) * sin(slope) + 0.607 * sin(folded_asp) * sin(slope))
}

samps$pdir <- pdir(samps$lat, samps$slope, samps$folded_aspect)
```

Consolidate the names of the topography variables
```{r consolidate_topo_vars}
topo_vars <- c("c_aspect", "pdir", "elev", "lat", "lon", "slope", "topo_roughness_1", "topo_roughness_2", "topo_roughness_3", "topo_roughness_4")
```

## Timing variables

Adjust ordinal day to reflect "middle of summer-ness" with August 6th (the 218th day of the year) getting a 1 and February 4th (the 35th day of the year) getting a value of -1

```{r adjust_ordinal_day}
circular_doy <- function(doy) {
  
  new_doy <- ((doy - 218) / 365) * 2 * pi
  return(cos(new_doy))
  
}

samps$c_doy <- circular_doy(samps$ordinal_day)
```

Make year an integer

```{r year_to_integer}
samps$year_ <- as.numeric(as.character(samps$year_))
```

Consolidate timing variable names.

```{r consolidate_timing_vars}
timing_vars <- c("alarm_year", "alarm_date", "cont_date", "c_doy")
```

## Vegetation variables

Consolidate vegetation variable names.

```{r consolidate_veg_var_names}
veg_vars <-
  sapply(X = 1:4, 
         FUN = function(i) paste(
           c(
             "focal_mean_ndvi",
             "focal_mean_ndwi",
             "het_ndvi",
             "het_ndwi",
             "ndvi_asm",
             "ndvi_contrast",
             "ndvi_corr",
             "ndvi_dent",
             "ndvi_diss",
             "ndvi_dvar",
             "ndvi_ent",
             "ndvi_idm",
             "ndvi_imcorr1",
             "ndvi_imcorr2",
             "ndvi_inertia",
             "ndvi_prom",
             "ndvi_savg",
             "ndvi_sent",
             "ndvi_shade",
             "ndvi_svar",
             "ndvi_var"
           ), i, sep = "_")) %>% 
  as.vector() %>% 
  c("preFire_ndvi", "preFire_ndwi")
```

## Fire weather variables

Consolidate fire weather variable names.

```{r consolidate_fire_weather_var_names}
fireWeather_vars <- c("erc", "fm100", "tmmx")
```

# Scale all predictor variables

First consolidate all variable names.

```{r consolidate_var_names}
all_vars <- c(topo_vars, timing_vars, veg_vars, fireWeather_vars)
```

Just focus on the points from the mixed conifer/yellow pine forest

```{r conifer_subset}
mixed_con <- 
  samps %>%
  filter(conifer_forest == 1)
```

Scale the predictors (ss = scaled samples)

```{r scale_samples}
ss <-
  mixed_con %>%
  mutate_at(.vars = all_vars, .funs = funs(s = as.numeric(scale(.))))

glimpse(ss)  
```

## High severity/non-high severity

```{r get_cbi_calibration_model_summary}
model_summary <- read.csv(here::here("data/data_output/cbi_calibration_model_comparison.csv"), stringsAsFactors = FALSE)
target_model <- model_summary[model_summary$response == "RBR" &
                                model_summary$time_window == 48 &
                                model_summary$interpolation == "bicubic", ]
```

Here are the other thresholds for the model using bicubic interpolation and a 48 day window:

- `r target_model$unchanged` equates to a CBI of 0
- `r target_model$low_sev` equates to a CBI of 0.1 -- threshold between "unchanged" and "low"
- `r target_model$mod_sev` equates to a CBI of 1.25 -- threshold between "low" and "medium"
- `r target_model$hi_sev` equates to a CBI of 2.25 -- threshold between "medium" and "high"

```{r high_v_not-high}
ss$stand_replacing <- ifelse(ss$RBR > target_model$hi_sev, yes = 1, no = 0)
```

Transform RBR to linear scale CBI

For 100-hour fuel moisture, we refer to @Stephens2012 who find 7.7% to be the 80th
percentile condition, 6.6% to be the 90th percentile condition, and 4.2% to be 
the 97.5th percentile condition. See also the note in @Stephens2013a about the significance
of 80th percentile conditions.

Create a variable for "extreme conditions" versus "non-extreme" conditions (with
respect to fuel moisture) and interact it with the heterogeneity variable.

```{r extreme_fm100}
# Extreme percentiles correspeond to 80th, 90th, and 97.5th percentiles of 100 hour
# fuel moisture
extreme_fm100_percentiles <- c(7.7, 6.6, 4.2)
ss$extreme80_fm100 <- ifelse(ss$fm100 < 7.7, yes = 1, no = 0)
ss$extreme90_fm100 <- ifelse(ss$fm100 < 6.6, yes = 1, no = 0)
ss$extreme97.5_fm100 <- ifelse(ss$fm100 < 4.2, yes = 1, no = 0)
```

Consider also subsetting to only places that burned to avoid any impact of unburned area that happens to be within the fire perimeter.

```{r conditional_on_burning}
ss_burned <- ss %>% filter(ss$RBR > target_model$low_sev)
```

```{r}
remoteSev_to_cbi <- function(data, response, a, b, c) {
  log((data[, response, drop = TRUE] - a) / b) / c
}

ss_burned$cbi <- remoteSev_to_cbi(data = ss_burned, response = "RBR", a = target_model$a, b = target_model$b, c = target_model$c)
```


## Modeling

```{r continuous_response}
m1a <- lm(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s, data = ss)
m2a <- lm(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s, data = ss)
m3a <- lm(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s, data = ss)
m4a <- lm(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s, data = ss)

summary(m1a)
summary(m2a)
summary(m3a)
summary(m4a)

m5a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m6a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m7a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m8a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m5a)
summary(m6a)
summary(m7a)
summary(m8a)

m9a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m10a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m11a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m12a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m9a)
summary(m10a)
summary(m11a)
summary(m12a)

m13a <- lmer(RBR ~ het_ndvi_1_s * fm100_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m14a <- lmer(RBR ~ het_ndvi_2_s * fm100_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m15a <- lmer(RBR ~ het_ndvi_3_s * fm100_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m16a <- lmer(RBR ~ het_ndvi_4_s * fm100_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m13a)
summary(m14a)
summary(m15a)
summary(m16a)

m17a <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m18a <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m19a <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m20a <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

summary(m17a)
summary(m18a)
summary(m19a)
summary(m20a)

AIC(m17a, m18a, m19a, m20a)

AIC(m1a, m2a, m3a, m4a, m5a, m6a, m7a, m8a, m9a, m10a, m11a, m12a, m13a, m14a, m15a, m16a, m17a, m18a, m19a, m20a) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)
```
```{r}
m1b <- lm(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s, data = ss_burned)
m2b <- lm(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s, data = ss_burned)
m3b <- lm(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s, data = ss_burned)
m4b <- lm(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s, data = ss_burned)

summary(m1a)
summary(m2a)
summary(m3a)
summary(m4a)

m5b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m6b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m7b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m8b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m5b)
summary(m6b)
summary(m7b)
summary(m8b)

m9b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m10b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m11b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m12b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m9b)
summary(m10b)
summary(m11b)
summary(m12b)

m13b <- lmer(RBR ~ het_ndvi_1_s * fm100_s + focal_mean_ndvi_1_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m14b <- lmer(RBR ~ het_ndvi_2_s * fm100_s + focal_mean_ndvi_2_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m15b <- lmer(RBR ~ het_ndvi_3_s * fm100_s + focal_mean_ndvi_3_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m16b <- lmer(RBR ~ het_ndvi_4_s * fm100_s + focal_mean_ndvi_4_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m13b)
summary(m14b)
summary(m15b)
summary(m16b)

m17b <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m18b <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m19b <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m20b <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

summary(m17b)
summary(m18b)
summary(m19b)
summary(m20b)

AIC(m17b, m18b, m19b, m20b)


AIC(m1b, m2b, m3b, m4b, m5b, m6b, m7b, m8b, m9b, m10b, m11b, m12b, m13b, m14b, m15b, m16b) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)
```


```{r binomial_response}

# Severe or not as bernoulli response, fm100 covariate, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_rawFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, fm100 covariate, only samples that were burned
m_sevOrNot_rawFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_rawFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

###

# Severe or not as bernoulli response, extreme or not dummy covariate, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_extremeFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, only samples that were burned
m_sevOrNot_extremeFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, all samples from within fire perimeter (some may not have burned)
m_sevOrNot_extremeFm100AndRawFm100_1_ss <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_2_ss <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_3_ss <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_4_ss <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss, control = glmerControl(optimizer = "bobyqa"))

# Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned
m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- glmer(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- glmer(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- glmer(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))
m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- glmer(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = "binomial", data = ss_burned, control = glmerControl(optimizer = "bobyqa"))

AIC(m_sevOrNot_rawFm100_1_ss, m_sevOrNot_rawFm100_2_ss, m_sevOrNot_rawFm100_3_ss, m_sevOrNot_rawFm100_4_ss,
    m_sevOrNot_extremeFm100_1_ss, m_sevOrNot_extremeFm100_2_ss, m_sevOrNot_extremeFm100_3_ss, m_sevOrNot_extremeFm100_4_ss,
    m_sevOrNot_extremeFm100AndRawFm100_1_ss, m_sevOrNot_extremeFm100AndRawFm100_2_ss, m_sevOrNot_extremeFm100AndRawFm100_3_ss, m_sevOrNot_extremeFm100AndRawFm100_4_ss)  %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)

AIC(m_sevOrNot_rawFm100_1_ssBurned, m_sevOrNot_rawFm100_2_ssBurned, m_sevOrNot_rawFm100_3_ssBurned, m_sevOrNot_rawFm100_4_ssBurned,
    m_sevOrNot_extremeFm100_1_ssBurned, m_sevOrNot_extremeFm100_2_ssBurned, m_sevOrNot_extremeFm100_3_ssBurned, m_sevOrNot_extremeFm100_4_ssBurned,
    m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)


# Best binomial response model; all samps
summary(m_sevOrNot_extremeFm100AndRawFm100_2_ss)

# Best binomial response model; burned samps
summary(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)


m_nonextreme <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = subset(ss_burned, subset = ss_burned$extreme80_fm100 == 0), control = glmerControl(optimizer = "bobyqa"))
m_extreme <- glmer(stand_replacing ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = "binomial", data = subset(ss_burned, subset = ss_burned$extreme80_fm100 == 1), control = glmerControl(optimizer = "bobyqa"))

summary(m_nonextreme)
summary(m_extreme)

```

```{r rethinking_models}
m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned <- 
  glmer(stand_replacing ~ 
          het_ndvi_1_s * 
          preFire_ndvi_s * 
          fm100_s * 
          focal_mean_ndvi_1_s + 
          topo_roughness_1_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

m_sevOrNot_meanNDVI_localNDVI_rawFm100_2_ssBurned <- 
  glmer(stand_replacing ~ 
          het_ndvi_2_s * 
          preFire_ndvi_s * 
          fm100_s * 
          focal_mean_ndvi_2_s +
          topo_roughness_2_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

m_sevOrNot_meanNDVI_localNDVI_rawFm100_3_ssBurned <- 
  glmer(stand_replacing ~ 
          het_ndvi_3_s * 
          preFire_ndvi_s * 
          fm100_s * 
          focal_mean_ndvi_3_s +
          topo_roughness_3_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

m_sevOrNot_meanNDVI_localNDVI_rawFm100_4_ssBurned <- 
  glmer(stand_replacing ~ 
          het_ndvi_4_s * 
          preFire_ndvi_s * 
          fm100_s * 
          focal_mean_ndvi_4_s + 
          topo_roughness_4_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

fm100_labeller <- c(`-0.5` = "Low fuel moisture", 
                    `0` = "Mean fuel moisture",
                    `0.5` = "High fuel moisture")

focal_ndvi_labeller <- c(`-2` = "Low neighborhood NDVI", 
                    `0` = "Mean neighborhood NDVI",
                    `2` = "High neighborhood NDVI")

e_1_4way <- effect("het_ndvi_1_s:preFire_ndvi_s:fm100_s:focal_mean_ndvi_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(preFire_ndvi_s = c(0, 0), fm100_s = c(-0.5, 0, 0.5), focal_mean_ndvi_1_s = c(-2, 0, 2), het_ndvi_1_s = seq(-2, 2, by = 0.1)))

e_1_4way_allEffects <- effect("het_ndvi_1_s:preFire_ndvi_s:fm100_s:focal_mean_ndvi_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(preFire_ndvi_s = seq(-2, 2, by = 0.25), fm100_s = seq(-2, 2, by = 0.25), focal_mean_ndvi_1_s = c(0, 0), het_ndvi_1_s = seq(-2, 2, by = 0.25)))

glimpse(as.data.frame(e_1_4way_allEffects))
e_1_4way_allEffects_gg <- as.data.frame(e_1_4way_allEffects)

# prefire NDVI main effect
prefireNDVI_ggplot <-
  ggplot(subset(e_1_4way_allEffects_gg, e_1_4way_allEffects_gg$fm100_s == 0 & e_1_4way_allEffects_gg$focal_mean_ndvi_1_s == 0 & e_1_4way_allEffects_gg$het_ndvi_1_s == 0), aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Prefire vegetation density") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))


# fm100 main effect
fm100_ggplot <-
  ggplot(subset(e_1_4way_allEffects_gg, e_1_4way_allEffects_gg$preFire_ndvi_s == 0 & e_1_4way_allEffects_gg$focal_mean_ndvi_1_s == 0 & e_1_4way_allEffects_gg$het_ndvi_1_s == 0), aes(x = fm100_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Fuel moisture") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))


pahl_effect <- effect("pdir_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(pdir_s = seq(-2, 2, by = 0.25)))
pahl_effect_gg <- as.data.frame(pahl_effect)

# pahl main effect
pahl_ggplot <- 
  ggplot(pahl_effect_gg, aes(x = pdir_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Annual heat load") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))



topo_rough_effect <- effect("topo_roughness_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(topo_roughness_1_s = seq(-2, 2, by = 0.25)))
topo_rough_effect_gg <- as.data.frame(topo_rough_effect)

# topo rough main effect
topo_rough_ggplot <-
  ggplot(topo_rough_effect_gg, aes(x = topo_roughness_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Topographic roughness") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

library(cowplot)

panel_plot <- plot_grid(NULL, prefireNDVI_ggplot, fm100_ggplot, NULL, pahl_ggplot, topo_rough_ggplot, ncol=3, nrow = 2, rel_widths = c(0.2, 1, 1))

panel_plot_labelled <- 
  panel_plot + 
  draw_label(label = "Probability of\n90%\ntree mortality", x = 0.08, angle = 0, size = 28)

ggplot2::ggsave(plot = panel_plot_labelled, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/usual-suspects.png", device = "png", width = 15)


e_1_4way_gg <- as.data.frame(e_1_4way)
e_1_4way_ggplot <- ggplot(e_1_4way_gg, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) + 
  geom_line(lwd = 1.25) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  facet_wrap(~ fm100_s, labeller = as_labeller(fm100_labeller)) +
  ylab(label = "Probability of >90% tree mortality") +
  xlab(label = "Neighborhood forest variability (scaled)\n(90m x 90m window)") +
  scale_color_viridis_d(name = "Neighborhood NDVI\n(scaled)") +
  scale_fill_viridis_d(name = "Neighborhood NDVI\n(scaled)") +
  theme_bw() +
  theme(text = element_text(size=28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

ggsave(plot = e_1_4way_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/4-way-plot.png", device = "png", width = 15, units = "in")
```

```{r simplify_models}
fm1 <- 
  glmer(stand_replacing ~ 
          het_ndvi_1_s +
          focal_mean_ndvi_1_s +
          het_ndvi_1_s:focal_mean_ndvi_1_s +
          preFire_ndvi_s +
          preFire_ndvi_s:focal_mean_ndvi_1_s +
          het_ndvi_1_s:preFire_ndvi_s +
          fm100_s +
          het_ndvi_1_s:fm100_s +
          topo_roughness_1_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

fm2 <- 
  glmer(stand_replacing ~ 
          het_ndvi_2_s +
          focal_mean_ndvi_2_s +
          het_ndvi_2_s:focal_mean_ndvi_2_s +
          preFire_ndvi_s +
          preFire_ndvi_s:focal_mean_ndvi_2_s +
          het_ndvi_2_s:preFire_ndvi_s +
          fm100_s +
          topo_roughness_2_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

fm3 <- 
  glmer(stand_replacing ~ 
          het_ndvi_3_s +
          focal_mean_ndvi_3_s +
          het_ndvi_3_s:focal_mean_ndvi_3_s +
          preFire_ndvi_s +
          preFire_ndvi_s:focal_mean_ndvi_3_s +
          het_ndvi_3_s:preFire_ndvi_s +
          fm100_s +
          topo_roughness_3_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

fm4 <- 
  glmer(stand_replacing ~ 
          het_ndvi_4_s +
          focal_mean_ndvi_4_s +
          het_ndvi_4_s:focal_mean_ndvi_4_s +
          preFire_ndvi_s +
          preFire_ndvi_s:focal_mean_ndvi_4_s +
          het_ndvi_4_s:preFire_ndvi_s +
          fm100_s +
          topo_roughness_4_s + 
          pdir_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

AIC(fm1, fm2, fm3, fm4, m_sevOrNot_meanNDVI_localNDVI_rawFm100_4_ssBurned)
summary(fm1)

e_fm1 <- allEffects(fm1, 
                    xlevels = list(preFire_ndvi_s = seq(-2, 2, length.out = 17), 
                                   fm100_s = seq(-2, 2, length.out = 17), 
                                   focal_mean_ndvi_1_s = seq(-2, 2, length.out = 17), 
                                   het_ndvi_1_s = seq(-2, 2, length.out = 17),
                                   topo_roughness_1_s = seq(-2, 2, length.out = 17),
                                   pdir_s = seq(-2, 2, length.out = 17)))

e_fm1_topoRough <- as.data.frame(e_fm1[["topo_roughness_1_s"]])
e_fm1_pahl <- as.data.frame(e_fm1[["pdir_s"]])
e_fm1_het_neighborhoodMean <- as.data.frame(e_fm1[["het_ndvi_1_s:focal_mean_ndvi_1_s"]])
e_fm1_focalMean_neighborhoodMean <- as.data.frame(e_fm1[["focal_mean_ndvi_1_s:preFire_ndvi_s"]])
e_fm1_het_focalMean <- as.data.frame(e_fm1[["het_ndvi_1_s:preFire_ndvi_s"]])
e_fm1_het_fm100 <- as.data.frame(e_fm1[["het_ndvi_1_s:fm100_s"]])

# PrefireNDVI_ggplot
e_fm1_het_focalMean_subset <- 
  e_fm1_het_focalMean %>% 
  filter(het_ndvi_1_s == 0)

prefireNDVI_ggplot <-
  ggplot(e_fm1_het_focalMean_subset, aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Prefire vegetation density") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# fm100 effect
e_fm1_fm100 <- 
  e_fm1_het_fm100 %>% 
  filter(het_ndvi_1_s == 0)

fm100_ggplot <- 
  ggplot(e_fm1_fm100, aes(x = fm100_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Fuel moisture") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# topo roughness effect
topo_rough_ggplot <- 
  ggplot(e_fm1_topoRough, aes(x = topo_roughness_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Topographic roughness") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# pahl effect
pahl_ggplot <- 
  ggplot(e_fm1_pahl, aes(x = pdir_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Annual heat load") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

panel_plot <- plot_grid(NULL, prefireNDVI_ggplot, fm100_ggplot, NULL, pahl_ggplot, topo_rough_ggplot, ncol=3, nrow = 2, rel_widths = c(0.2, 1, 1))

panel_plot_labelled <- 
  panel_plot + 
  draw_label(label = "Probability of\n90%\ntree mortality", x = 0.08, angle = 0, size = 28)

ggplot2::ggsave(plot = panel_plot_labelled, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/usual-suspects.png", device = "png", width = 15)

# heterogeneity interacting with fuel moisture effect (not quite significant)
e_fm1_het_fm100_subset <-
  e_fm1_het_fm100 %>% 
  filter(fm100_s %in% c(-2, 0, 2))
         
het_fm100_ggplot <- 
  ggplot(e_fm1_het_fm100_subset, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(fm100_s), fill = as.factor(fm100_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Neighborhood variability (scaled)") +
  ylab(label = "") +
  scale_color_viridis_d(name = "Fuel moisture (scaled)") +
  scale_fill_viridis_d(name = "Fuel moisture (scaled)") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# heterogeneity interacting with neighborhood mean effect (not significant)
ggplot(e_fm1_het_neighborhoodMean, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

# focal mean interacting with neighborhood mean effect (strong interaction)
e_fm1_focalMean_neighborhoodMean_subset <-
  e_fm1_focalMean_neighborhoodMean %>% 
  filter(focal_mean_ndvi_1_s %in% c(-2, 0, 2))

e_fm1_focalMean_neighborhoodMean_ggplot <-
  ggplot(e_fm1_focalMean_neighborhoodMean_subset, aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d(name = "Neighborhood NDVI (scaled)") +
  scale_color_viridis_d(name = "Neighborhood NDVI (scaled)") +
  xlab(label = "Prefire vegetation density (scaled)") +
  ylab(label = "Probability of 90% tree mortality") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# focal mean interacting with neighborhood mean effect (strong interaction)
e_fm1_focalMean_neighborhoodMean_subset_2 <-
  e_fm1_focalMean_neighborhoodMean %>% 
  filter(preFire_ndvi_s %in% c(-2, 0, 2))

e_fm1_focalMean_neighborhoodMean_ggplot_2 <-
  ggplot(e_fm1_focalMean_neighborhoodMean_subset_2, aes(x = focal_mean_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(preFire_ndvi_s), fill = as.factor(preFire_ndvi_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d(name = "Prefire NDVI (scaled)") +
  scale_color_viridis_d(name = "Prefire NDVI (scaled)") +
  xlab(label = "Neighborhood NDVI (scaled)") +
  ylab(label = "Probability of 90% tree mortality") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

ggplot2::ggsave(plot = e_fm1_focalMean_neighborhoodMean_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/focalMean_neighborhoodMean.png", device = "png")

# heterogeneity interacting with focal mean effect (not significant)
ggplot(e_fm1_het_focalMean, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(preFire_ndvi_s), fill = as.factor(preFire_ndvi_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

# Heterogeneity main effect
het_effect_gg <- 
  e_fm1_het_focalMean %>% 
  filter(preFire_ndvi_s == 0)

het_effect_ggplot <-
  ggplot(het_effect_gg, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  ylab(label = "Probability of 90% tree mortality") +
  xlab(label = "Neighborhood forest variability (scaled)\n(90m x 90m window)") +
  theme(text = element_text(size = 28),
        axis.text.x = element_text(size = 28, margin = margin(t = 14, b = 14, l = 0, r = 0)),
        axis.text.y = element_text(size = 28, margin = margin(t = 0, b = 0, l = 14, r = 14)),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

ggplot2::ggsave(plot = het_effect_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/het_effect.png", device = "png")
```

```{r rbr_response}

# RBR response, fm100 covariate, all samples from within fire perimeter (some may not have burned)
m_rbr_rawFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_rawFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, fm100 covariate, only samples that were burned
m_rbr_rawFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_rawFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

###

# RBR response, extreme or not dummy covariate, all samples from within fire perimeter (some may not have burned)
m_rbr_extremeFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, extreme or not dummy covariate, only samples that were burned
m_rbr_extremeFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

# RBR response, extreme or not dummy covariate, interaction with raw fm100, all samples from within fire perimeter (some may not have burned)
m_rbr_extremeFm100AndRawFm100_1_ss <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_2_ss <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_3_ss <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss)
m_rbr_extremeFm100AndRawFm100_4_ss <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss)

# RBR response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned
m_rbr_extremeFm100AndRawFm100_1_ssBurned <- lmer(RBR ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_2_ssBurned <- lmer(RBR ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_3_ssBurned <- lmer(RBR ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), data = ss_burned)
m_rbr_extremeFm100AndRawFm100_4_ssBurned <- lmer(RBR ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), data = ss_burned)

AIC(m_rbr_rawFm100_1_ss, m_rbr_rawFm100_2_ss, m_rbr_rawFm100_3_ss, m_rbr_rawFm100_4_ss,
    m_rbr_extremeFm100_1_ss, m_rbr_extremeFm100_2_ss, m_rbr_extremeFm100_3_ss, m_rbr_extremeFm100_4_ss,
    m_rbr_extremeFm100AndRawFm100_1_ss, m_rbr_extremeFm100AndRawFm100_2_ss, m_rbr_extremeFm100AndRawFm100_3_ss, m_rbr_extremeFm100AndRawFm100_4_ss)  %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)

AIC(m_rbr_rawFm100_1_ssBurned, m_rbr_rawFm100_2_ssBurned, m_rbr_rawFm100_3_ssBurned, m_rbr_rawFm100_4_ssBurned,
    m_rbr_extremeFm100_1_ssBurned, m_rbr_extremeFm100_2_ssBurned, m_rbr_extremeFm100_3_ssBurned, m_rbr_extremeFm100_4_ssBurned,
    m_rbr_extremeFm100AndRawFm100_1_ssBurned, m_rbr_extremeFm100AndRawFm100_2_ssBurned, m_rbr_extremeFm100AndRawFm100_3_ssBurned, m_rbr_extremeFm100AndRawFm100_4_ssBurned) %>% 
  mutate(model = rownames(.)) %>% 
  arrange(AIC)


# All samps best model
summary(m_rbr_rawFm100_4_ss)

# Only burned best model
summary(m_rbr_extremeFm100_4_ssBurned)

# Model to match best model from stand replacing binomial response
summary(m_rbr_extremeFm100AndRawFm100_1_ss)
summary(m_rbr_extremeFm100AndRawFm100_2_ss)
summary(m_rbr_extremeFm100AndRawFm100_3_ss)
summary(m_rbr_extremeFm100AndRawFm100_4_ss)

```

<!-- ```{r} -->
<!-- # Calculating the differences between B11, B12, and B13: -->
<!-- H0s <- rbind(het_normal = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0), -->
<!--              het_xtreme = c(0, 1, 1, 0, 0, 0, 0, 1, 0, 0, 0), -->
<!--              norm_xtreme = c(0, 0, -1, 0, 0, 0, 0, -1, 0, 0, 0)) -->

<!-- # Binomial response, all samps -->
<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_1_ss, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ss) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_2_ss, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ss) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_3_ss, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ss) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ss <- glht(m_sevOrNot_extremeFm100AndRawFm100_4_ss, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ss) -->

<!-- # Binomial response, burned samps -->
<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned) -->

<!-- tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- glht(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned) -->

<!-- # RBR response, all samps -->
<!-- tH0s_m_rbr_extremeFm100AndRawFm100_1_ss <- glht(m_rbr_extremeFm100AndRawFm100_1_ss, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_1_ss) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_2_ss <- glht(m_rbr_extremeFm100AndRawFm100_2_ss, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_2_ss) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_3_ss <- glht(m_rbr_extremeFm100AndRawFm100_3_ss, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_3_ss) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_4_ss <- glht(m_rbr_extremeFm100AndRawFm100_4_ss, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_4_ss) -->

<!-- # RBR response, burned samps -->
<!-- tH0s_m_rbr_extremeFm100AndRawFm100_1_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_1_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_1_ssBurned) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_2_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_2_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_2_ssBurned) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_3_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_3_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_3_ssBurned) -->

<!-- tH0s_m_rbr_extremeFm100AndRawFm100_4_ssBurned <- glht(m_rbr_extremeFm100AndRawFm100_4_ssBurned, linfct = H0s) -->
<!-- summary(tH0s_m_rbr_extremeFm100AndRawFm100_4_ssBurned) -->

<!-- ``` -->

<!-- ```{r fit_with_brms} -->
<!-- library(brms) -->
<!-- # Severe or not as bernoulli response, extreme or not dummy covariate, interaction with raw fm100, only samples that were burned -->

<!-- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_1_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_1_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned) -->

<!-- m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_2_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_2_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned) -->

<!-- m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_3_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_3_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned) -->

<!-- m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm <- brm(stand_replacing ~ het_ndvi_4_s * extreme80_fm100 * fm100_s + preFire_ndvi_s + topo_roughness_4_s + pdir_s + (1 | fire_id), family = bernoulli(), data = ss_burned) -->

<!-- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm) -->
<!-- # m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm) -->
<!-- # m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm) -->
<!-- # m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm <- brms::add_loo(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm) -->
<!-- ``` -->

<!-- Set up the proper contrasts for each model -->
<!-- ```{r contrasts_setup} -->
<!-- # Contrasts -->
<!-- # Fix these so that fm100_s isn't 0, but rather the mean for each group (greater than 0 for non-extreme, less than 0 for extreme) -->

<!-- # What is the mean fm100_s for non-extreme and extreme conditions? -->
<!-- fm100_s_means <-  -->
<!--   ss_burned %>%  -->
<!--   as.data.frame() %>%  -->
<!--   group_by(extreme80_fm100) %>%  -->
<!--   summarize(mean_fm100 = mean(fm100_s)) -->

<!-- het_normal <- c(0, 1, 0, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 0), mean_fm100)), 0, 0, 0, 0, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 0), mean_fm100)), 0, 0) -->

<!-- het_xtreme <- c(0, 1, 1, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100)) -->
<!-- , 0, 0, 0, 1, pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100)) -->
<!-- , pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100)) -->
<!-- , pull(dplyr::select(filter(fm100_s_means, extreme80_fm100 == 1), mean_fm100)) -->
<!-- ) -->
<!-- ``` -->

<!-- Look at smallest neighborhood effect of heterogeneity: -->

<!-- ```{r one_pixel_neighbornood} -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm) -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_samps[, 1:11] -->

<!-- het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas, 1, function(x) x %*% het_normal) -->
<!-- het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_1_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned)(0) -->

<!-- hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned, breaks = 100) -->
<!-- ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_1_ssBurned)(0) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned - het_xtreme_samps, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_1_ssBurned - het_xtreme_samps)(0) -->
<!-- ``` -->

<!-- ```{r two_pixel_neighbornood} -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm) -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_samps[, 1:11] -->

<!-- het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas, 1, function(x) x %*% het_normal) -->
<!-- het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_2_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned)(0) -->

<!-- hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned, breaks = 100) -->
<!-- ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_2_ssBurned)(0) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned - het_xtreme_samps, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_2_ssBurned - het_xtreme_samps)(0) -->
<!-- ``` -->

<!-- ```{r three_pixel_neighbornood} -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm) -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_samps[, 1:11] -->

<!-- het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas, 1, function(x) x %*% het_normal) -->
<!-- het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_3_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned)(0) -->

<!-- hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned, breaks = 100) -->
<!-- ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_3_ssBurned)(0) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned - het_xtreme_samps, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_3_ssBurned - het_xtreme_samps)(0) -->
<!-- ``` -->

<!-- ```{r four_pixel_neighbornood} -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_samps <- posterior_samples(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm) -->
<!-- m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas <- m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_samps[, 1:11] -->

<!-- het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas, 1, function(x) x %*% het_normal) -->
<!-- het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned <- apply(m_sevOrNot_extremeFm100AndRawFm100_4_ssBurned_brm_betas, 1, function(x) x %*% het_xtreme) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)(0) -->

<!-- hist(het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned, breaks = 100) -->
<!-- ecdf(het_xtreme_sevOrNot_extremeFm100AndRawFm100_4_ssBurned)(0) -->

<!-- hist(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned - het_xtreme_samps, breaks = 100) -->
<!-- ecdf(het_normal_sevOrNot_extremeFm100AndRawFm100_4_ssBurned - het_xtreme_samps)(0) -->
<!-- ``` -->


