---
title: "ESA2018 presentation"
author: "Michael Koontz"
date: "9/25/2018"
output: html_document
---

```{r setup, message = FALSE}
library(lubridate)
library(sf)
library(lme4)
library(dplyr)
library(ggplot2)
library(here)
library(lubridate)
library(multcomp)
library(effects)
library(cowplot)
```

```{r read_files, message = FALSE}
fires <- st_read(dsn = here::here("/data/data_output/fire_perim/fire17_1_sn"),
                 stringsAsFactors = FALSE) %>% 
  st_transform(4326)

sn <- st_read(here::here("/data/features/SierraEcoregion_Jepson/SierraEcoregion_Jepson.shp")) %>% 
  st_transform(4326)

all_or_burned <- "burned"

if (all_or_burned == "all") {
  if (!file.exists(here::here("data/data_output/all-fire-samples_texture_configured.rds"))) {
    source(here::here("data/data_carpentry/configure_fire-samples.R"))
  }
  load(here::here("data/data_output/all-fire-samples_texture_configured.rds"))
}

if (all_or_burned == "burned") {
  if (!file.exists(here::here("data/data_output/burned-fire-samples_texture_configured.rds"))) {
      source(here::here("data/data_carpentry/configure_fire-samples.R"))
    }
    load(here::here("data/data_output/burned-fire-samples_texture_configured.rds"))
}

```

```{r models}
m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned <- 
  glmer(stand_replacing ~ 
          het_ndvi_1_s * 
          preFire_ndvi_s * 
          fm100_s * 
          focal_mean_ndvi_1_s + 
          topo_roughness_1_s + 
          pahl_s + 
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))

fm1 <- 
  glmer(stand_replacing ~ 
          het_ndvi_1_s +
          focal_mean_ndvi_1_s +
          preFire_ndvi_s +
          fm100_s +
          pahl_s + 
          topo_roughness_1_s + 
          focal_mean_ndvi_1_s:preFire_ndvi_s +
          het_ndvi_1_s:focal_mean_ndvi_1_s +
          het_ndvi_1_s:preFire_ndvi_s +
          het_ndvi_1_s:fm100_s +
          (1 | fire_id), 
        family = "binomial", 
        data = ss_burned, 
        control = glmerControl(optimizer = "bobyqa"))```
```


```{r effects}
fm100_labeller <- c(`-0.5` = "Low fuel moisture", 
                    `0` = "Mean fuel moisture",
                    `0.5` = "High fuel moisture")

focal_ndvi_labeller <- c(`-2` = "Low neighborhood NDVI", 
                    `0` = "Mean neighborhood NDVI",
                    `2` = "High neighborhood NDVI")

e_1_4way <- effect("het_ndvi_1_s:preFire_ndvi_s:fm100_s:focal_mean_ndvi_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(preFire_ndvi_s = c(0, 0), fm100_s = c(-0.5, 0, 0.5), focal_mean_ndvi_1_s = c(-2, 0, 2), het_ndvi_1_s = seq(-2, 2, by = 0.1)))

e_1_4way_allEffects <- effect("het_ndvi_1_s:preFire_ndvi_s:fm100_s:focal_mean_ndvi_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(preFire_ndvi_s = seq(-2, 2, by = 0.25), fm100_s = seq(-2, 2, by = 0.25), focal_mean_ndvi_1_s = c(0, 0), het_ndvi_1_s = seq(-2, 2, by = 0.25)))

glimpse(as.data.frame(e_1_4way_allEffects))
e_1_4way_allEffects_gg <- as.data.frame(e_1_4way_allEffects)

# prefire NDVI main effect
prefireNDVI_ggplot <-
  ggplot(subset(e_1_4way_allEffects_gg, e_1_4way_allEffects_gg$fm100_s == 0 & e_1_4way_allEffects_gg$focal_mean_ndvi_1_s == 0 & e_1_4way_allEffects_gg$het_ndvi_1_s == 0), aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Prefire vegetation density") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))


# fm100 main effect
fm100_ggplot <-
  ggplot(subset(e_1_4way_allEffects_gg, e_1_4way_allEffects_gg$preFire_ndvi_s == 0 & e_1_4way_allEffects_gg$focal_mean_ndvi_1_s == 0 & e_1_4way_allEffects_gg$het_ndvi_1_s == 0), aes(x = fm100_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Fuel moisture") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))


pahl_effect <- effect("pahl_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(pahl_s = seq(-2, 2, by = 0.25)))
pahl_effect_gg <- as.data.frame(pahl_effect)

# pahl main effect
pahl_ggplot <- 
  ggplot(pahl_effect_gg, aes(x = pahl_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Annual heat load") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))



topo_rough_effect <- effect("topo_roughness_1_s", m_sevOrNot_meanNDVI_localNDVI_rawFm100_1_ssBurned, xlevels = list(topo_roughness_1_s = seq(-2, 2, by = 0.25)))
topo_rough_effect_gg <- as.data.frame(topo_rough_effect)

# topo rough main effect
topo_rough_ggplot <-
  ggplot(topo_rough_effect_gg, aes(x = topo_roughness_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Topographic roughness") +
  ylab(label = "") +
  theme_bw() +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

panel_plot <- plot_grid(NULL, prefireNDVI_ggplot, fm100_ggplot, NULL, pahl_ggplot, topo_rough_ggplot, ncol=3, nrow = 2, rel_widths = c(0.2, 1, 1))

panel_plot_labelled <- 
  panel_plot + 
  draw_label(label = "Probability of\n90%\ntree mortality", x = 0.08, angle = 0, size = 28)

ggplot2::ggsave(plot = panel_plot_labelled, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/usual-suspects.png", device = "png", width = 15)


e_1_4way_gg <- as.data.frame(e_1_4way)
e_1_4way_ggplot <- ggplot(e_1_4way_gg, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) + 
  geom_line(lwd = 1.25) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  facet_wrap(~ fm100_s, labeller = as_labeller(fm100_labeller)) +
  ylab(label = "Probability of >90% tree mortality") +
  xlab(label = "Neighborhood forest variability (scaled)\n(90m x 90m window)") +
  scale_color_viridis_d(name = "Neighborhood NDVI\n(scaled)") +
  scale_fill_viridis_d(name = "Neighborhood NDVI\n(scaled)") +
  theme_bw() +
  theme(text = element_text(size=28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

ggsave(plot = e_1_4way_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/4-way-plot.png", device = "png", width = 15, units = "in")
```

```{r esa-new-orleans}
e_fm1 <- allEffects(fm1, 
                    xlevels = list(preFire_ndvi_s = seq(-2, 2, length.out = 17), 
                                   fm100_s = seq(-2, 2, length.out = 17), 
                                   focal_mean_ndvi_1_s = seq(-2, 2, length.out = 17), 
                                   het_ndvi_1_s = seq(-2, 2, length.out = 17),
                                   topo_roughness_1_s = seq(-2, 2, length.out = 17),
                                   pahl_s = seq(-2, 2, length.out = 17)))

e_fm1_topoRough <- as.data.frame(e_fm1[["topo_roughness_1_s"]])
e_fm1_pahl <- as.data.frame(e_fm1[["pahl_s"]])
e_fm1_het_neighborhoodMean <- as.data.frame(e_fm1[["het_ndvi_1_s:focal_mean_ndvi_1_s"]])
e_fm1_focalMean_neighborhoodMean <- as.data.frame(e_fm1[["focal_mean_ndvi_1_s:preFire_ndvi_s"]])
e_fm1_het_focalMean <- as.data.frame(e_fm1[["het_ndvi_1_s:preFire_ndvi_s"]])
e_fm1_het_fm100 <- as.data.frame(e_fm1[["het_ndvi_1_s:fm100_s"]])

# PrefireNDVI_ggplot
e_fm1_het_focalMean_subset <- 
  e_fm1_het_focalMean %>% 
  filter(het_ndvi_1_s == 0)

prefireNDVI_ggplot <-
  ggplot(e_fm1_het_focalMean_subset, aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Prefire vegetation density") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# fm100 effect
e_fm1_fm100 <- 
  e_fm1_het_fm100 %>% 
  filter(het_ndvi_1_s == 0)

fm100_ggplot <- 
  ggplot(e_fm1_fm100, aes(x = fm100_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Fuel moisture") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# topo roughness effect
topo_rough_ggplot <- 
  ggplot(e_fm1_topoRough, aes(x = topo_roughness_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Topographic roughness") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# pahl effect
pahl_ggplot <- 
  ggplot(e_fm1_pahl, aes(x = pahl_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Annual heat load") +
  ylab(label = "") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

panel_plot <- plot_grid(NULL, prefireNDVI_ggplot, fm100_ggplot, NULL, pahl_ggplot, topo_rough_ggplot, ncol=3, nrow = 2, rel_widths = c(0.2, 1, 1))

panel_plot_labelled <- 
  panel_plot + 
  draw_label(label = "Probability of\n90%\ntree mortality", x = 0.08, angle = 0, size = 28)

# ggplot2::ggsave(plot = panel_plot_labelled, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/usual-suspects.png", device = "png", width = 15)

# heterogeneity interacting with fuel moisture effect (not quite significant)
e_fm1_het_fm100_subset <-
  e_fm1_het_fm100 %>% 
  filter(fm100_s %in% c(-2, 0, 2))
         
het_fm100_ggplot <- 
  ggplot(e_fm1_het_fm100_subset, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(fm100_s), fill = as.factor(fm100_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  xlab(label = "Neighborhood variability (scaled)") +
  ylab(label = "") +
  scale_color_viridis_d(name = "Fuel moisture (scaled)") +
  scale_fill_viridis_d(name = "Fuel moisture (scaled)") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# heterogeneity interacting with neighborhood mean effect (not significant)
ggplot(e_fm1_het_neighborhoodMean, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

# focal mean interacting with neighborhood mean effect (strong interaction)
e_fm1_focalMean_neighborhoodMean_subset <-
  e_fm1_focalMean_neighborhoodMean %>% 
  filter(focal_mean_ndvi_1_s %in% c(-2, 0, 2))

e_fm1_focalMean_neighborhoodMean_ggplot <-
  ggplot(e_fm1_focalMean_neighborhoodMean_subset, aes(x = preFire_ndvi_s, y = fit, ymin = lower, ymax = upper, color = as.factor(focal_mean_ndvi_1_s), fill = as.factor(focal_mean_ndvi_1_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d(name = "Neighborhood NDVI (scaled)") +
  scale_color_viridis_d(name = "Neighborhood NDVI (scaled)") +
  xlab(label = "Prefire vegetation density (scaled)") +
  ylab(label = "Probability of 90% tree mortality") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# focal mean interacting with neighborhood mean effect (strong interaction)
e_fm1_focalMean_neighborhoodMean_subset_2 <-
  e_fm1_focalMean_neighborhoodMean %>% 
  filter(preFire_ndvi_s %in% c(-2, 0, 2))

e_fm1_focalMean_neighborhoodMean_ggplot_2 <-
  ggplot(e_fm1_focalMean_neighborhoodMean_subset_2, aes(x = focal_mean_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(preFire_ndvi_s), fill = as.factor(preFire_ndvi_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d(name = "Prefire NDVI (scaled)") +
  scale_color_viridis_d(name = "Prefire NDVI (scaled)") +
  xlab(label = "Neighborhood NDVI (scaled)") +
  ylab(label = "Probability of 90% tree mortality") +
  theme(text = element_text(size = 28),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)),
        axis.text.x = element_text(size = 28),
        axis.text.y = element_text(size = 28))

# ggplot2::ggsave(plot = e_fm1_focalMean_neighborhoodMean_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/focalMean_neighborhoodMean.png", device = "png")

# heterogeneity interacting with focal mean effect (not significant)
ggplot(e_fm1_het_focalMean, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper, color = as.factor(preFire_ndvi_s), fill = as.factor(preFire_ndvi_s))) +
  geom_line(lwd = 1.25) + 
  geom_ribbon(alpha = 0.4) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

# Heterogeneity main effect
het_effect_gg <- 
  e_fm1_het_focalMean %>% 
  filter(preFire_ndvi_s == 0)

het_effect_ggplot <-
  ggplot(het_effect_gg, aes(x = het_ndvi_1_s, y = fit, ymin = lower, ymax = upper)) +
  geom_line(lwd = 1.25) +
  geom_ribbon(alpha = 0.4) +
  ylab(label = "Probability of 90% tree mortality") +
  xlab(label = "Neighborhood forest variability (scaled)\n(90m x 90m window)") +
  theme(text = element_text(size = 28),
        axis.text.x = element_text(size = 28, margin = margin(t = 14, b = 14, l = 0, r = 0)),
        axis.text.y = element_text(size = 28, margin = margin(t = 0, b = 0, l = 14, r = 14)),
        axis.title.x = element_text(margin = margin(t = 28, b = 28, l = 0, r = 0)),
        axis.title.y = element_text(margin = margin(t = 0, b = 0, l = 28, r = 28)))

# ggplot2::ggsave(plot = het_effect_ggplot, filename = "~/dev/presentations/2018-08-09_esa2018-new-orleans-louisiana/het_effect.png", device = "png")
```