---
title: "Explore texture via CBI plots"
author: "Michael Koontz"
date: "3/11/2018"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
library(ggplot2)
library(rlang)
library(sf)
library(dplyr)
library(mgcv)
library(purrr)
library(broom)
library(modelr)
library(tidyr)
library(lazyeval)
library(lme4)
library(here)
```

Load the data.

```{r load_data}
# Get 48-day window, bicubic interpolation data

cbi <- st_read(here::here("data/ee_cbi-calibration/cbi-calibration_48-day-window_L57_bicubic-interp.geojson"), stringsAsFactors = FALSE)

glimpse(cbi)
```

Some exploratory plots using a few predictors.

# Standard deviation of NDVI

Our original metric. Tells us about the spread of the distribution of NDVI on the same scale as that of the NDVI data

```{r het_ndvi}
hetNDVI <- cbi %>%
  select(starts_with("het_ndvi"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("het_ndvi")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(hetNDVI, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(hetNDVI, aes(x = heterogeneity, y = cbi_over)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))
```

# Standard deviation of NDWI 

```{r het_ndwi}
hetNDWI <- cbi %>%
  select(starts_with("het_ndwi"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("het_ndwi")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(hetNDWI, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(hetNDWI, aes(x = heterogeneity, y = cbi_over)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))
```

# Entropy

"Measures the randomness of a gray-level distribution"

```{r entropy}
nd_ent <- cbi %>% 
  select(starts_with("nd_ent"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_ent")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_ent, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_ent, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))
```

# Inverse difference moment

"Homogeneity. IDM tells us about the smoothness of the image."

```{r nd_idm}
nd_idm <- cbi %>% 
  select(starts_with("nd_idm"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = homogeneity, starts_with("nd_idm")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_idm, aes(x = homogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_idm, aes(x = homogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# Variance

"Variance tells us how spread out the distribution of gray-levels is."

```{r nd_var}
nd_var <- cbi %>% 
  select(starts_with("nd_var"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_var")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_var, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_var, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))
```

# Angular second moment

"ASM measures the number of repeated pairs"

```{r asm}
nd_asm <- cbi %>% 
  select(starts_with("nd_asm"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_asm")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_asm, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_asm, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# Correlation

"Measures the correlation between the two pixels in the pixel pair"

```{r nd_corr}
nd_corr <- cbi %>% 
  select(starts_with("nd_corr"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_corr")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_corr, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_corr, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# Contrast

"Contrast measures the local contrast of an image"

```{r nd_contrast}
nd_contrast <- cbi %>% 
  select(starts_with("nd_contrast"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_contrast")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_contrast, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_contrast, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```


# savg

"sum average"

```{r nd_savg}
savg <- cbi %>% 
  select(starts_with("nd_savg"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_savg")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(savg, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(savg, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# svar

"sum variance"

```{r nd_svar}
svar <- cbi %>% 
  select(starts_with("nd_svar"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_svar")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(svar, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(svar, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# sent

"sum entropy"

```{r nd_sent}
sent <- cbi %>% 
  select(starts_with("nd_sent"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_sent")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(sent, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(sent, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# dvar

"difference variance"

```{r dvar}
dvar <- cbi %>% 
  select(starts_with("nd_dvar"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_dvar")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(dvar, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(dvar, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# dent

"difference entropy"

```{r nd_dent}
dent <- cbi %>% 
  select(starts_with("nd_dent"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_dent")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(dent, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(dent, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# imcorr1

"Info. measure of corr. 1"

```{r nd_imcorr1}
imcorr1 <- cbi %>% 
  select(starts_with("nd_imcorr1"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_imcorr1")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(imcorr1, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(imcorr1, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```


# imcorr2

"Info. measure of corr. 2"

```{r nd_imcorr2}
imcorr2 <- cbi %>% 
  select(starts_with("nd_imcorr2"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_imcorr2")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(imcorr2, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(imcorr2, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# dissimilarity

"dissimilarity"

```{r dissimilarity}
nd_diss <- cbi %>% 
  select(starts_with("nd_diss"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_diss")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_diss, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_diss, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# inertia

"intertia"

```{r nd_inertia}
nd_inertia <- cbi %>% 
  select(starts_with("nd_inertia"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_inertia")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_inertia, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_inertia, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```

# shade

"cluster shade"

```{r nd_shade}
nd_shade <- cbi %>% 
  select(starts_with("nd_shade"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_shade")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_shade, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_shade, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```


# Prominence

"cluster prominence"

```{r nd_prom}
nd_prom <- cbi %>% 
  select(starts_with("nd_prom"), cbi_over, id, RBR) %>% 
  gather(key = radius, value = heterogeneity, starts_with("nd_prom")) %>% 
  separate(col = radius, into = c("type", "vi", "radius")) %>% 
  tidyr::unite(het_type, type, vi, sep = "_")

ggplot(nd_prom, aes(x = heterogeneity)) +
  geom_density() + 
  facet_wrap( ~ as.factor(radius))

ggplot(nd_prom, aes(x = heterogeneity, y = RBR)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_wrap( ~ as.factor(radius))

```